Semi-supervised low-rank logistic regression for high-dimensional neuroimaging data

### 2015-05-31 paper_nips/paper.tex
Comments on abstract and introduction

\begin{abstract}
Imaging neuroscience links human behavior to aspects of brain
biology in ever-increasing datasets.
%
Existing neuroimaging methods typically perform either discovery of unknown
neural structure or testing of neural structure associated with mental tasks.
%
However, testing hypotheses on the neural correlates of mental tasks
depends on the representation of the observations.
%
We therefore propose to optimize
representation and task classification in
a unified statistical learning problem.
%
A multinomial logistic regression is introduced that is
constrained by joint rank with a representation-enforcing
autoencoder.
%
We show that this approach yields more accurate and more interpretable
neural models of psychological tasks in a reference dataset.
%

@@

\section{Introduction}
%
Methods for neuroimaging research can be grouped by discovering
neurobiological structure or assessing the neural correlates associated
with mental tasks.
To discover, on the one hand, spatial distributions of neural activity
structure across time,
independent component analysis (ICA) \cite{beckmann2005} is often used.
It decomposes the BOLD (blood-oxygen level-dependent) signals into the
primary modes of variation.
The ensuing spatial activity patterns are believed to represent
brain networks of
functionally interacting regions \cite{smith2009}.
Similarly, sparse principal component analysis (SPCA) \cite{varoqu2011}
has been used to
separate BOLD signals into parsimonious network components.
Thus extracted brain networks are probably
manifestations of electrophysiological oscillation frequencies \cite{hipp15}.
Their fundamental organizational role is further
attested by continued covariation during sleep and anesthesia \cite{fox07}.
%
Network discovery by applying ICA or SPCA is typically performed on
task-unrelated (i.e., \textit{unlabeled}) ``resting-state'' data.
These capture brain dynamics
during ongoing random thought without controlled environmental stimulation.
In fact, a large proportion of the BOLD signal variation
is known
not to correlate with a particular behavior, stimulus, or experimental task
\cite{fox07}.

@@

Both these biological properties can be conjointly exploited in an
approach that is
mixed (i.e., using task and rest data)
and 
low-rank (i.e., performing network decomposition).
%
The integration of brain-network discovery into
supervised classification can yield a semi-supervised learning
framework. The most relevant
neurobiological structure should hence be identified
for the prediction problem at hand.
%
Autoencoders suggest themselves because they can emulate
variants of most unsupervised learning algorithms,
including PCA, SPCA, and ICA \cite{hinton06}.
\begin{wrapfigure}{r}{0.40\textwidth}
  \centering
  %\captionsetup{labelformat=nonumber}
    \includegraphics[width=0.40\textwidth]{figures/figure1.png}
  \vspace{-0.7cm}
  \caption {\textbf{Architecture.}
  Linear autoencoders find an optimal compression of 79,941 brain voxels
  into $n$ unknown activity patterns by improving reconstruction from them.
  The decomposition matrix equates with the bottleneck of
  a factored logistic regression.
  Supervised multi-class learning on task data ($X_{task}$)
  is thus guided by
  unsupervised decomposition of rest data ($X_{rest}$).
  }
\end{wrapfigure}
%
Autoencoders
are often one-layered learning models 
that condense the input data to
local and global 
representations
by optimizing reconstruction from them.
%
% XXX Michael: In fact, they behave like a non-centered PCA, 
% XXX which amounts to just doing an SVD
% XXX which btw is NOT a convex optimization problem, although easy to 
% XXX calculate and with a global optimum.
% XXX Autoencoders are also not one-layered by default. Let's make sure that
% XXX we don't trip up by getting the technicalities wrong and getting 
% XXX reviewers annoyed at us because of it.
% DanilO: cool thanks. That is the stuff, we need.
They behave like a (truncated, non-centered) PCA
in case of one linear hidden layer and a squared error loss
\cite{baldi1989neural}.
This architecture yields an optimization objective
with unique global minimum.
Autoencoders behave like a SPCA if shrinkage terms are added to the
model weights in the optimization objective.
Moreover, they have the characteristics of an ICA in case of tied weights
and adding a nonlinear convex function
at the first layer \cite{le2011ica}.
% XXX What convex transformation? Why is that relevant? Convexity is only
% XXX relevant for the loss
% Danilo: it's about adding a neural-network-like nonconvex function for
% first-layer activation
These authors further demonstrated that ICA, sparse autoencoders, and 
sparse coding are mathematically equivalent
under mild conditions.
%
Hence, autoencoders can flexibly project the neuroimaging data
onto the main axes of variation and thus % XXX Michael <- that's PCA
reverse-engineer properties of the data-generating
processes in the brain \cite{olshausen96}.
% XXX Michael ^ I don't get that last part nor the citation.
% Danilo: that is the magic

@@

>>>>

\begin{abstract}
Imaging neuroscience links human behavior to aspects of brain
biology in ever-increasing datasets.
%
Existing neuroimaging methods typically perform either discovery of unknown
neural structure or testing of neural structure associated with mental tasks.
%
However, testing hypotheses on the neural correlates of an increasing number of mental tasks makes it crucial to represent the observations adequately.
%
We therefore propose to optimize
representation and task classification in
a unified statistical learning problem.
%
A multinomial logistic regression is introduced that is
constrained by joint rank with a representation-enforcing
autoencoder.
%
We show that this approach yields more accurate and more interpretable
neural models of psychological tasks in a reference dataset, and better generalization to other datasets.
%

@@

\section{Introduction}
%
Methods for neuroimaging research can be grouped by discovering
neurobiological structure or assessing the neural correlates associated
with mental tasks.
To discover, on the one hand, spatial distributions of neural activity
structure across time,
independent component analysis (ICA) \cite{beckmann2005} is often used.
It decomposes the BOLD (blood-oxygen level-dependent) signals into the
primary modes of variation.
The ensuing spatial activity patterns are believed to represent
brain networks of
functionally interacting regions \cite{smith2009}.
Similarly, sparse principal component analysis (SPCA) \cite{varoqu2011}
has been used to
separate BOLD signals into parsimonious network components.
The extracted brain networks are probably
manifestations of electrophysiological oscillation frequencies \cite{hipp15}.
Their fundamental organizational role is further
attested by continued covariation during sleep and anesthesia \cite{fox07}.
%
Network discovery by applying ICA or SPCA is typically performed on
task-unrelated (i.e., \textit{unlabeled}) ``resting-state'' data.
These capture brain dynamics
during ongoing random thought without controlled environmental stimulation.
In fact, a large proportion of the BOLD signal variation
is known
not to correlate with a particular behavior, stimulus, or experimental task
\cite{fox07}.

@@

Both these biological properties can be conjointly exploited in an
approach that is
mixed (i.e., using task and rest data)
and 
low-rank (i.e., performing network decomposition).
%
The integration of brain-network discovery into
supervised classification can yield a semi-supervised learning
framework. The most relevant
neurobiological structure should hence be identified
for the prediction problem at hand.
%
Autoencoders suggest themselves because they can emulate
variants of most unsupervised learning algorithms,
including PCA, SPCA, and ICA \cite{hinton06}.
\begin{wrapfigure}{r}{0.40\textwidth}
  \centering
  %\captionsetup{labelformat=nonumber}
    \includegraphics[width=0.40\textwidth]{figures/figure1.png}
  \vspace{-0.7cm}
  \caption {\textbf{Architecture.}
  Linear autoencoders find an optimal compression of 79,941 brain voxels
  into $n$ unknown activity patterns by improving reconstruction from them.
  The decomposition matrix equates with the bottleneck of
  a factored logistic regression.
  Supervised multi-class learning on task data ($X_{task}$)
  is thus guided by
  unsupervised decomposition of rest data ($X_{rest}$).
  }
\end{wrapfigure}
%
Autoencoders
are often one-layered learning models 
that condense the input data to
local and global 
representations
by optimizing reconstruction from them.
%
% XXX Michael: In fact, they behave like a non-centered PCA, 
% XXX which amounts to just doing an SVD
% XXX which btw is NOT a convex optimization problem, although easy to 
% XXX calculate and with a global optimum.
% XXX Autoencoders are also not one-layered by default. Let's make sure that
% XXX we don't trip up by getting the technicalities wrong and getting 
% XXX reviewers annoyed at us because of it.
% DanilO: cool thanks. That is the stuff, we need.
They behave like a (truncated, non-centered) PCA
in case of one linear hidden layer and a squared error loss
\cite{baldi1989neural}.
This architecture yields an optimization objective
with unique global minimum.
Autoencoders behave like a SPCA if shrinkage terms are added to the
model weights in the optimization objective.
Moreover, they have the characteristics of an ICA in case of tied weights
and adding a nonlinear convex function
at the first layer \cite{le2011ica}.
% XXX What convex transformation? Why is that relevant? Convexity is only
% XXX relevant for the loss
% Danilo: it's about adding a neural-network-like nonconvex function for
% first-layer activation
These authors further demonstrated that ICA, sparse autoencoders, and 
sparse coding are mathematically equivalent
under mild conditions.
%
Hence, autoencoders can flexibly project the neuroimaging data
onto the main axes of variation and thus % XXX Michael <- that's PCA
reverse-engineer properties of the data-generating
processes in the brain \cite{olshausen96}.
% XXX Michael ^ I don't get that last part nor the citation.
\textbf{XXX Bertrand: neither do I. I think you exxagerate.}
% Danilo: that is the magic

@@

\textbf{XXX Bertrand: I think that you could emphasize the problem that we are always data-poor, so that the set of representations that  can be extracted from SPMs taken from  a few tens of subjects is limited. What you propose it to extend it by i) co-analysing many problems simultaneously (multi-task) and using unlabled data that spana space of meaningful configurations
}
\textbf{XXX Bertrand: We need a sentence like: Our contribution is 1. XXX 2.YYY etc.}
